---
title: "Counting the number of placements of queens on chessboard"
output: html_notebook
---

This notebook contains two simple functions poscount and queenchess.

1. poscount: Counts the total number of possible placements of n queens on an nxn chessboard given that no queen can attack the other in one turn. This function can also return the list of all possible layouts.

2. queenchess: This function plots the possible placements of queens on a chessboard. Maximum number of boards that can be displayed is bounded above by 20 as the default.

```{r count}
poscount <- function(n, j0 = 1, mat, givelayout = FALSE, qlays = list()){
  
  if(j0 > n){
    if(givelayout){
      qlays[[length(qlays) + 1]] = mat
      return(list(qlays = qlays, s = 1))
    }else{
      return(list(s = 1))
    }
  }
  
  if(missing(mat))
    mat = data.frame(x = numeric(0), y = numeric(0))
  
  if(nrow(mat) == n){
    if(givelayout){
      qlays[[length(qlays) + 1]] = mat
      return(list(qlays = qlays, s = 1))
    }else{
      return(list(s = 1))
    }
  }
  
  sum1 = 0
  
  sett = setdiff(1:n, c(unique(mat$y), j0 + mat$y - mat$x, mat$y + mat$x - j0))
  
  if(length(sett) == 0){
    if(givelayout){
      return(list(qlays = qlays, s = 0))
    }else{
      return(list(s = 0))
    }
  }
  
  for(k0 in sett){
    mat0 = mat
    mat0 = rbind.data.frame(mat0, data.frame(x = j0, y = k0))
    if(givelayout){
      pp = poscount(n, j0 + 1, mat = mat0, givelayout = givelayout, qlays = qlays)
      sum1 = sum1 + pp$s
      qlays = pp$qlays
    }else{
      sum1 = sum1 + poscount(n, j0 + 1, mat = mat0)$s
    }
  }
  if(givelayout){
    return(list(qlays = qlays, s = sum1))
  }else{
    return(list(s = sum1))
  }
}
```

```{r plot function}
queenchess <- function(qlays, maxboard = 20, randomorder = TRUE, tsize){
  
  mat1 = data.frame()
  n = max(qlays[[1]]$x)
  if(length(qlays) > maxboard){
    if(randomorder){
      iset = sample(1:length(qlays), maxboard)  
    }else{
      iset = 1:maxboard
    }
  }else{
    iset = 1:length(qlays)
  }
  for(i in iset){
  
  mat = as.data.frame(do.call("rbind", lapply(1:n, function(i) cbind(i, 1:n))))
  colnames(mat) = c('x', 'y')
  qq = qlays[[i]]
  qq$room = 1
  qq$piece = sprintf("\u265b")

  mat = merge(mat, qq, all.x = TRUE, by = c('x', 'y'))
  mat$i = i
  mat1 = rbind.data.frame(mat1, mat)
  }
  
  if(missing(tsize))
    tsize = ceiling(24 - length(iset))
  
  ggplot(mat1, aes(x = x, y = y, fill = (x+y)%%2)) +
    theme(legend.position = "none", axis.ticks = element_blank(), panel.background = element_blank(),
          strip.background = element_blank(), strip.text.x = element_blank(),
          axis.text.x = element_blank(), axis.text.y = element_blank(), aspect.ratio = 1) +
    geom_tile(color = "white", size = 1) +
    xlab(NULL) + ylab(NULL) +
    geom_text(aes(label = piece, color = (x+y+1)%%2), size = tsize) +
    facet_wrap(~i)
}
```
```{r counts}
cat("The number of possible layouts for an 8x8 and a 10x10chessboard are ", poscount(n = 8)$s, " and ", poscount(n= 10)$s, '.\n', sep = '')
```


Below we can see two example plots. The size of the queen can be adjusted depending on plot size.
```{r plot, display = TRUE, warning=FALSE}
plot(queenchess(poscount(n = 8, givelayout = TRUE)$qlays, maxboard = 6, tsize = 5))
plot(queenchess(poscount(n = 5, givelayout = TRUE)$qlays, tsize = 5))
plot(queenchess(poscount(n = 4, givelayout = TRUE)$qlays))
```
